\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-09 trunk/arch.tex
updates

The \rx architecture was designed with the objective to bring planning
closer to the low-level control of the vehicle while maintaining its
reactivity. Indeed, in most of the architecture where task planning is
embedded, this task remains at a failry high level of abstraction as
the complexity of this task would otherwise impede the system
reactiveness. As a result the planning problem managed {\em in situ}
remains fairly detached from low-level control and most of the
reactive adaptation which in contrast is managed by specialized
component that relies on different representations for modeling the
local behavior they control or even do not have any explicit formal
model of this behavior. The two angles our architecture attck this
issue are the following:

@@

For the first point \rx provides a formal framework described in
details in \cite{py10} and \cite{rajan12} and is not the focus of this
paper. The second have deep implication on how on should integrate a
planner in this framework in general and was applied to the europa
framework specifically. We'll develop later how this requirement
implies to see in-situ planning as two parallels processes that are
tightly linked through the same plan structure. But before that we
need to introduce at least the high leve lconcepts and ideas behind
our architecture as a whole.

\rx architecture is structured around the notion of {\em state
  variables} as the unique basis of interactionalong with a well
defined ownership model for each of these.  In this framework each
reactor is seen as a black box providing its own {\em internal} state
variables for which its is reponsible to maintain its value at every
single instant -- represented by a {\em tick} with a fixed
duration. In order to identify its state a reactor can subscribe to
{\em external} state variables for which it will received new updates
when the owner of these will produce its {\em internal} state. This
{\em synchronization} process progate through the reactor dependency
graph on a bottom up flow at every {\em tick} ensuring that all the
reactors have a consistent view of the world. Conversely each reactor
can request a future change on one of its {\em external} state
variable which will be transfromed as an {\em internal} goal to the
owner. This goal is expected to porduce a deliberation phase on the
receiver which can in turn produce a plan impacting its {\em external}
state variables that will propagate down following the same mechanism.

@@

Take for example the \rx agent instance we use in our AUV
(Fig. \ref{fig:agent}). In this architecture all the reactors
(symbolized by colored boxes) do not have explicit connection from one
to another but instead rely on a publish/subscribe model on state
variables. For Example the \textsf{Pilot} reactor needs the {\em
  externally} managed state informations from \textsf{Sensor\_Data}
and \textsf{Nav\_cmd} states variables in order to maintain {\em
  internally} the state variables \textsf{Waypoint} and
\textsf{Nav\_policy}. This relation applies in two way :

@@

\begin{figure}[!htb]
 \centering
 \includegraphics[scale=0.45]{figs/AUV-agent.pdf}
 \caption{\small A \texttt{T-REX} agent is composed of multiple
   reactors or control loops (rounded boxes) which are connected
   through state variables provided by one reactor ({\color{blue}blue}
   solid line) with multiple possible clients ({\color{green}green}
   dashed lines)}
  \label{fig:agent}
\end{figure}

Overall the \rx architecture helps to abstract to the reactor
interaction to only state variables. Each reactor is practically
ignoring where its {\em external} state information is managed as long
as this information is available and properly maintained so they can
properly manage their own {\em internal} state information which in
turn may be used by other reactors. Such formalism also constrain
reactors interaction to state information exchanges which are two
folds:

@@

\item {\em Observation on the current state values}: This information
  propagates up in the reactor graph at the synchronization phase 
  and occur at the beginning of every time steps (called
  ticks). During this phase the agent provides to each reactor 
  updates on their {\em external} state variables for this tick and
  ask them to compute their {\em internal} state variables values for
  this same tick. By doing so we propagate a consistent view of that
  state of the world throughout all the reactors as time advance.

\item {\em Goals request on future state}: This information propagates
  down in the reactor graph. It reflects that, in order to satisfy own
  of its {\em internal} objectives, a reactor may have a plan that
  rely on future values of one of its {\em external} state. When such
  a plan is identified we ensure that this part of the plan is given
  as a goal to the owner of the given {\em external(s)} state variable(s).

@@

The choice of representation of state variables, observations and
goals is directly derieved to the timleines and tokens from europa.
such representation is ich enough to allow the excahnge of state
information in a formal and yet flexible way and its transaltion in
planning frameworks that have an explicit representation of time (such
as but not limited to europa) is often trivial.

The overall execution cycle of the agent is also abstracted as some
continuous planning/deliberation cycle between all the
reactors. Indeed , at the agent interface level, each reactor provided
only two abstract calls that conceptualize their execution:

@@

The way the agent interleave planning \texttt{step}s and
synchronization is illustrated in Figure \ref{fig:tick-exec}.
Synchronization happens at every single tick and will interrupt
planning allowing this reactor to identify its current state that can
propagate through the reactor graph in order to ensure t aconsitent
view of the state of the world between all the reactors.

@@

This view of the reactor has two parallel tasks aloow for long
planning deliberation while ensuring that the state evolution
propagates through the agent as time advance. These two tasks 
intertwined together give the two natural flows (bottom-up for state
estimation and top to bottom for future plan) that we expect to see in
a control loop. As they are both conceptualized as inference processes
it is natural to implement a reactor based on generic automated
planning. For example in Fig. \ref{fig:agent}, appart \textsf{Shore
  comm.} and \textsf{Vehicle} acting as interfaces to the system,
every other reactors (colored in blue) are different instance of our
europa reactor differntiated only by their model and europa solver
configurations.

@@

The \eu reactor is our implementation inside \rx of a reactor that
usess automated planning for its execution. This reactor is fully
based on the europa framework and applies the core idea we of \rx we
described above to implement a reactor for which the sole purpse is to
deliberate as time advance. The choice of europa as the planning
framework was directed by the fact thta this framework provide a rich
and highly configurable framework to be able to infer around a rich
temporal domain.

@@

This assumption, while helpfull to reduce the scope of the planning
problem, becomes problematic if one wants to embed a planner in a
real-time system that evolves in a highly dynamic environment. In a
typical robot this implies that the only point where the planner can
be informed about the current world state is righe before the planning
is started by giving to the planner the intial state and current
objectives. During the planning, it is assumed that the agent can
maintin the world as perceived by the planner as stable. In
\cite{lemai04, lemai-chenevier2004}, they do attempt to reduce the
impact of planning by aloowing local plan repairs which attempt to
insert repairing actions in the exisiting plan andf can be done fast
enough. Still in ythe eventuality that such repair is not possible
they add to stop their vehicle until the full (re-)planning was
complete. {\em I may need more examples/refs here}

>>>>

\subsection{Introductory concepts}
\label{sec:arch:intro}

A prime motivation behind the \rx architecture was to design a
controller which could bring automated planning closer to low-level
control of the vehicle while maintaining its reactivity. Traditionally
planning has been deemed computationally expensive with planner
performance restricting robot reactivity
\cite{ghallab04,Dias:2003ua}. In control architectures where task
planning is embedded, planning consequently remains at an abstract
level with the assumption that it would otherwise impede system
reactiveness and that the environment can change at a faster rate than
the planner can plan for it. In such a situation, the agent may thrash
if the internal state of the plan gets out of synch with the actual
state of the world. As a result the planning problem managed {\em in
  situ} remains fairly detached from low-level control. Most
adaptations at the lower-level reactive layers in contrast are managed
by specialized components (for example GeNoM \comment{} and CLARATy
\comment{}) that rely on different representations for modeling local
control behavior or worse do not have any explicit formal agent
behavior model. The consequent different techniques for specifying
each layer in the architecture results in duplication of effort and a
diffusion of knowledge leading to significant design and integration
issues \cite{DS1report}.  \texttt{IDEA} mitigated most of these issues
by tightly integrating planning and execution in a single
representational and computational framework with a unified
declarative model. Our approach with \texttt{T-REX} builds on this
design methodology albeit in a substantially systematic manner

The \rx architecture contributes to agent architectures over and
beyond \texttt{IDEA} are in the following ways:

@@

\rx provides a formal framework which details partitioning in
\cite{py10} and \cite{rajan12} and is not the focus of this
chapter. Coupled interaction between planning and the world on the
other hand, has a significant implication on how planning can be
integrated in such frameworks; here we show using the \eu planner can
be coupled to our framework. First we link concepts in \eu planning
with the \rx framework. Note that in the following treatise, we call
the formal semantic framework of \rx as the 'agent'. Reactors
constitute an entity within this agent and are standalone blocks of
inference.

% on should integrate a
% planner in this framework in general and was applied to the europa
% framework specifically. We'll develop later how this requirement
% implies to see in-situ planning as two parallels processes that are
% tightly linked through the same plan structure. But before that we
% need to introduce at least the high leve lconcepts and ideas behind
% our architecture as a whole.

\subsection{Architectural concepts in \rx}
\label{sec:arch:trex}

The \rx architecture is structured around the notion of \eu {\em state
  variables} as the fundamental basis of interaction. Additionally it
comes with a well-defined ownership model associated with each state
variable. In this framework each reactor is seen as a 'black box'
providing its own {\em internal} state variables which it is
responsible for in order to maintain its value at every instant of
time. Each such instance is represented by a {\em tick} with a fixed
duration; in our implementation of \rx on the Dorado, we use a tick of
$1$ second. In order to identify its state a reactor can subscribe to
{\em external} state variables owned by other reactors, for which it
will receive new updates when the owner reactor generates its {\em
  internal} state.

This {\em synchronization} process propagates up through the reactor
dependency graph bottom up at every {\em tick} ensuring that all the
reactors have a consistent view of the world. Conversely each reactor
can request a future change on one of its {\em external} state
variables which will be transfromed as an {\em internal} goal to the
owner. Such \emph{goal dispatching} is expected to produce a
deliberation phase for the receiving reactor which can cyclically and
in turn, produce a plan impacting its {\em external} state variables
that will propagate down following the same mechanism.

@@

\begin{figure}[!htb]
 \centering
 \includegraphics[scale=0.45]{figs/AUV-agent.pdf}
 \caption{\small A \rx agent is composed of multiple reactors or
   control loops (rounded boxes) which are connected through state
   variables provided by one reactor ({\color{blue}blue} solid line)
   with multiple possible clients ({\color{green}green} dashed lines)}
  \label{fig:agent}
\end{figure}

Consider for example the \rx agent instance we use on our AUV as shown
in Fig. \ref{fig:agent}. In this instance all the reactors (symbolized
by colored boxes) do not have an explicit connection from one to
another but instead rely on a publish/subscribe model on shared state
variables. For instance the \textsf{Pilot} reactor needs the {\em
  externally} managed state information from \textsf{Sensor\_Data} and
\textsf{Nav\_cmd} state variables in order to maintain the
\emph{internal} state variables \textsf{Waypoint} and
\textsf{Nav\_policy}. This relation applies in two ways:

@@

Given the above example, the \rx architecture therefore helps abstract
reactor interaction by containing them to state variables. Each
reactor is therefore agnostic where its {\em external} state
information is managed as long as this information is available and
properly maintained for managing {\em internal} state
information. This {\em internal} state in turn may be used by other
reactors. Such a formalism also constrains reactor interaction to
state information exchange which comes in two forms:

@@

\item {\em Observation on the current state values}: This information
  propagates \emph{up} in the reactor dependency graph at the
  synchronization phase and occurs at the beginning of every
  tick. During this phase the agent provides updates to each reactor
  on their {\em external} state variables for this tick so as they
  could compute their {\em internal} state variable values for this
  same tick. By doing so we propagate a consistent view of that state
  of the world throughout all the reactors as time
  advances. \comment{could do with a figure here}

\item {\em Goals request on future state}: This information propagates
  \emph{down} in the reactor dependency graph. In order to satisfy own
  {\em internal} objectives, a reactor may have a plan that relies on
  future values of one of its {\em external} state. When such a plan
  is identified the agent ensures that this part of the plan is given
  as a goal to the owner of the given {\em external} state
  variable(s). \comment{could do with a figure here}

@@

The choice of representation of state variables, observations and
goals is directly derieved to the timelines and tokens from europa. 
such representation is rich enough to allow the exchange of state
information in a formal and yet flexible way and its transaltion in
planning frameworks that have an explicit representation of time (such
as but not limited to europa) is often trivial.

The overall execution cycle of the agent is also abstracted as a
continuous planning/deliberation cycle between all the
reactors. At the agent interface level, each reactor provided
only two abstract calls that conceptualize their execution:

@@

The way the agent interleave planning \texttt{step}s and
synchronization for a single reactor is illustrated in
Fig.~\ref{fig:tick-exec}. Synchronization happens at every single tick
and will interrupt planning allowing this reactor to identify its
current state that can propagate through the reactor graph in order to
ensure t aconsitent view of the state of the world between all the
reactors.

@@

This view of the reactor has two parallel tasks allow for long
planning deliberation while ensuring that the state evolution
propagates through the agent as time advance. These two tasks 
intertwined together give the two natural flows (bottom-up for state
estimation and top to bottom for future plan) that we expect to see in
a control loop. As they are both conceptualized as inference processes
it is natural to implement a reactor based on generic automated
planning. For example in Fig. \ref{fig:agent}, appart \textsf{Shore
  comm.} and \textsf{Vehicle} acting as interfaces to the system,
every other reactors (colored in blue) are different instance of our
europa reactor differntiated only by their model and europa solver
configurations.

@@

The \eu reactor is our implementation inside \rx of a reactor that
usess automated planning for its execution. This reactor is fully
based on the europa framework and applies the core idea of \rx we
described above to implement a reactor for which the sole purpose is
to deliberate as time advance. The design of this reactor takes cues
from the IDEA architecture presented on \cite{mus02, mus06} in the
sense that both planning and execution are based on a single unified
model. The choice of \eu as the planning framework was directed by the
fact that this framework provide a rich and highly configurable
toolset to be able to infer around a rich temporal domain.

@@

This assumption, while helpfull to reduce the scope of the planning
problem, becomes problematic if one wants to embed a planner in a
real-time system that evolves in a highly dynamic environment. In a
typical robot this implies that the only point where the planner can
be informed about the current world state is right before the planning
is started by giving to the planner the intial state and current
objectives. During the planning, it is assumed that the agent can
maintin the world as perceived by the planner as stable. In
\cite{lemai04, lemai-chenevier2004}, they do attempt to reduce the
impact of planning by allowing local plan repairs which attempt to
insert repairing actions in the exisiting plan andf can be done fast
enough. Still in ythe eventuality that such repair is not possible
they add to stop their vehicle until the full (re-)planning was
complete. {\em I may need more examples/refs here}
